### 2020-11-29 test.Rmd
test.rmd upload, for models update 11/29

---
title: "TEST"
output: pdf_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lmtest)
library(sandwich)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(patchwork)
library(stargazer)
library(tinytex)
library(dplyr)
library(car)
```

```{r variables, warning=FALSE, results='hide'}
df <- read.csv("covid-19.csv", header = TRUE)
df<-df%>%
  rename(case_rate_100k = 'Case.Rate.per.100000',
         death_rate_100k = 'Death.Rate.per.100000',
         population_density = 'Population.density.per.square.miles',
         white_pct = 'White...of.Cases',
         black_pct = 'Black...of.Cases',
         hispanic_pct = 'Hispanic...of.Cases',
         other_pct = 'Other...of.Cases',
         state_emergency = 'State.of.emergency',
         business_closed = 'Closed.other.non.essential.businesses',
         business_reopen = 'Began.to.reopen.businesses.statewide',
         mask_public='Mandate.face.mask.use.by.all.individuals.in.public.spaces',
         mask_legal='No.legal.enforcement.of.face.mask.mandate',
         black_population_pct ="Black...of.Total.Population",
         white_population_pct = "White...of.Total.Population",
         poverty_pct = "Percent.living.under.the.federal.poverty.line..2018.",
         unemployed_pct = "Percent.Unemployed..2018.",
         senior_pct = "X65.",
         population_2018 = "Population.2018", ## extra variable in test
         non_elderly_pre_existing = "Nonelderly.Adults.Who.Have.A.Pre.Existing.Condition",
         homeless = "Number.Homeless..2019."
         ) %>%
  select(State, case_rate_100k, death_rate_100k,population_density, white_pct, black_pct, hispanic_pct, 
         other_pct, state_emergency, business_closed, business_reopen, mask_public, mask_legal, 
         black_population_pct, white_population_pct, poverty_pct, unemployed_pct, senior_pct,
         population_2018, non_elderly_pre_existing, homeless)## extra variable in test
# Assign correct data types
cols.num <- c("white_pct","black_pct","hispanic_pct","other_pct")
df[cols.num] <- sapply(df[cols.num],as.numeric)
df$state_emergency=as.Date(df$state_emergency, format = "%m/%d/%Y")
df$business_closed=as.Date(df$business_closed, format = "%m/%d/%Y")
df$business_reopen=as.Date(df$business_reopen, format = "%m/%d/%Y")
head(df)
```
"Average social distance": average area of the square shape each people occupies, the length of such square is assumed to the average social distance(meter)
1 square miles = 2589988.11 square meters
```{r}
df<-df %>%
  mutate(
    homeless_pct = homeless/population_2018,
    noneld_pre_existing_pct = non_elderly_pre_existing/population_2018,
    avg_distance = sqrt(2589988.11/population_density),
    area_per_capita = 2589911.11/population_density
  )
```

```{r}
ggplot(data = df,
  mapping = aes(x= homeless_pct))+
  geom_histogram(fill = 'skyblue', color = 'grey30', breaks=seq(0,0.05,0.001)) +
  labs(title =  "Homeless percentage", x = "Homeless percentage", 
       y = 'Count')#+
  #scale_x_continuous(breaks=seq(0.1, 0.25, 0.05)
```
```{r}
ggplot(data = df,
  mapping = aes(x= noneld_pre_existing_pct))+
  geom_histogram(fill = 'skyblue', color = 'grey30', breaks=seq(0.1,0.25,0.01)) +
  labs(title =  "Non-elderly pre-existing condition percentage", x = "percentage", y = 'Count')
  #+scale_x_continuous(breaks=seq(0.1, 0.25, 0.05))
```
```{r}
hist_log_distance <- ggplot(data = df,
  mapping = aes(x= log(avg_distance)))+
  geom_histogram(fill = 'skyblue', color = 'grey30') +
  labs(title =  "log of Average distabce", x = "log of Meter", 
       y = 'Count')

hist_distance <- ggplot(data = df,
  mapping = aes(x= avg_distance))+
  geom_histogram(fill = 'skyblue', color = 'grey30') +
  labs(title =  "Average distabce", x = "Meter", 
       y = 'Count')

hist_log_distance/hist_distance
```
```{r}
hist_log_area <- ggplot(data = df,
  mapping = aes(x= log(area_per_capita)))+
  geom_histogram(fill = 'skyblue', color = 'grey30') +
  labs(title =  "log of Average area", x = "log of Meter^2", 
       y = 'Count')

hist_area <- ggplot(data = df,
  mapping = aes(x= area_per_capita))+
  geom_histogram(fill = 'skyblue', color = 'grey30') +
  labs(title =  "Average area", x = "Meter^2", 
       y = 'Count')

hist_log_area/hist_area
```
```{r}
model_test <- lm(case_rate_100k ~ log(avg_distance) + poverty_pct + noneld_pre_existing_pct + senior_pct + unemployed_pct+ white_population_pct, data = df)
coeftest(model_test)
vif(model_test)
plot(model_test, which = 3)
```

```{r}
model_test_b <- lm(case_rate_100k ~ log(population_density) + poverty_pct + noneld_pre_existing_pct + senior_pct + unemployed_pct+ white_population_pct, data = df)
coeftest(model_test_b)
vif(model_test_b)
plot(model_test_b, which = 3)
```
```{r}
model_test_b <- lm(case_rate_100k ~   poverty_pct + senior_pct + unemployed_pct, data = df)
coeftest(model_test_b, vcov = vcovHC)
vif(model_test_b)
plot(model_test_b, which = 3)
```
```{r}
model_test_b <- lm(case_rate_100k ~   noneld_pre_existing_pct+poverty_pct + senior_pct + unemployed_pct, data = df)
coeftest(model_test_b, vcov = vcovHC)
vif(model_test_b)
plot(model_test_b, which = 3)
```