### 2020-12-01 Model_2_final_ver.Rmd
model_2 _final, regression & clm assumptions

---
title: "Model_2_final_version"
output: pdf_document
---
```{r}
install.packages("GGally")
```

 **Initial reassignment of common data used across models:**
```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lmtest)
library(sandwich)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(patchwork)
library(stargazer)
library(tinytex)
library(dplyr)
library(car)
```

```{r variables, warning=FALSE, results='hide'}
df <- read.csv("covid-19.csv", header = TRUE)
df<-df%>%
  rename(case_rate_100k = 'Case.Rate.per.100000',
         death_rate_100k = 'Death.Rate.per.100000',
         population_density = 'Population.density.per.square.miles',
         white_pct = 'White...of.Cases',
         black_pct = 'Black...of.Cases',
         hispanic_pct = 'Hispanic...of.Cases',
         other_pct = 'Other...of.Cases',
         state_emergency = 'State.of.emergency',
         business_closed = 'Closed.other.non.essential.businesses',
         business_reopen = 'Began.to.reopen.businesses.statewide',
         mask_public='Mandate.face.mask.use.by.all.individuals.in.public.spaces',
         mask_legal='No.legal.enforcement.of.face.mask.mandate',
         black_population_pct ="Black...of.Total.Population",
         white_population_pct = "White...of.Total.Population",
         poverty_pct = "Percent.living.under.the.federal.poverty.line..2018.",
         unemployed_pct = "Percent.Unemployed..2018.",
         senior_pct = "X65.",
         population_2018 = "Population.2018", ## extra variable in test
         non_elderly_pre_existing = "Nonelderly.Adults.Who.Have.A.Pre.Existing.Condition",
         homeless = "Number.Homeless..2019."
         ) %>%
  select(State, case_rate_100k, death_rate_100k,population_density, white_pct, black_pct, hispanic_pct, 
         other_pct, state_emergency, business_closed, business_reopen, mask_public, mask_legal, 
         black_population_pct, white_population_pct, poverty_pct, unemployed_pct, senior_pct,
         population_2018, non_elderly_pre_existing, homeless)## extra variable in test
# Assign correct data types
cols.num <- c("white_pct","black_pct","hispanic_pct","other_pct")
df[cols.num] <- sapply(df[cols.num],as.numeric)
df$state_emergency=as.Date(df$state_emergency, format = "%m/%d/%Y")
df$business_closed=as.Date(df$business_closed, format = "%m/%d/%Y")
df$business_reopen=as.Date(df$business_reopen, format = "%m/%d/%Y")
head(df)
```

```{r}
df<-df %>%
  mutate(
    noneld_pre_existing_per_100k = 100000*non_elderly_pre_existing/population_2018,
    white_per_100k = 100000*white_population_pct,
    senior_per_100k = 100000*senior_pct,
    poverty_per_100k = 100000*poverty_pct, 
    unemployed_per_100k = 100000*unemployed_pct,
    log_population_density = log(population_density),
    avg_distance = sqrt(2589988.11/population_density),
    area_per_capita = 2589911.11/population_density
  )
```
## Second Model

### Model Regression

Case Rate per 100000 = senior_pct + nonelderly + poverty_pct + population_density + white_pct

$$
\begin{aligned}
 Case.Rate.Per.100K =\beta_0 &+\beta_1Senior.Rate.Per.100K \\&+\beta_2Nonelderly.Pre.Existing.Rate.Per.100K\\&+\beta_3Poverty.Rate.Per.100K+\beta_4ln(Population.Density)\\&+\beta_5White.Rate.Per.100K
\end{aligned}
$$
```{r}
model_2 <- lm(case_rate_100k ~ senior_per_100k + log(population_density) + noneld_pre_existing_per_100k + poverty_per_100k + white_per_100k, data = df)
coeftest(model_2)
```

## Limits
###i.IID Sampling

###ii.Linear Conditional Expectation
Assessing non-linear in higher-dimensional space: look at the predicted vs. residuals of the model.

```{r, echo=FALSE}
df<- df%>% 
  mutate(
    model_2_preds = predict(model_2), 
    model_2_resids = resid(model_2)
  ) 
df %>% 
  ggplot(aes(model_2_preds, model_2_resids)) + 
  geom_point() + 
  stat_smooth()
```

###iii.No Perfect Collinearity

First, check if any variables were dropped by R. No coefficient is missing, which mean there is no perfect collinearity.

```{r}
model_2$coefficients
```
In addition, all the variance inflation factors are less than 4, which doesn't indicate the exisitence of collinearity. 
```{r}
vif(model_2)
```
###iv.Homoskedastic Errors
32 is New Mexico, 33 is New York, 35 is North Dakota

```{r, echo=FALSE}
plot(model_2, which=3)
```
###v.Normally Distributed Errors
Check the normality of error distribution based on the histogram of residuals and the qqplot.

```{r, echo=FALSE}
plot_one <- df %>% 
  ggplot(aes(x = model_2_resids)) + 
  stat_bin(bins = 25) +
  geom_histogram()

plot_two <- df %>% 
  ggplot(aes(sample = model_2_resids)) + 
  stat_qq() + stat_qq_line()

plot_one / plot_two
```
### Cook Distance 
9-DC, 12-Hawaii, 32-New Mexico.

```{r, echo =FALSE}
plot(model_2, which = 5)
```

$$
\begin{aligned}
 Case.Rate.Per.100K =\beta_0 &+\beta_1Senior.Rate.Per.100K \\&+\beta_2Nonelderly.Pre.Existing.Rate.Per.100K\\&+\beta_3Poverty.Rate.Per.100K+\beta_4ln(Population.Density)\\&+\beta_5White.Rate.Per.100K+?\beta_6Unemployed.per.100k
\end{aligned}
$$
```{r}
model_2b <- lm(case_rate_100k ~ senior_per_100k + log(population_density) + noneld_pre_existing_per_100k + poverty_per_100k + white_per_100k + unemployed_per_100k, data = df)
coeftest(model_2b)
```
## i.IID Sampling

###ii.Linear Conditional Expectation
Assessing non-linear in higher-dimensional space: look at the predicted vs. residuals of the model.

```{r, echo=FALSE}
df<- df%>% 
  mutate(
    model_2b_preds = predict(model_2b), 
    model_2b_resids = resid(model_2b)
  ) 
df %>% 
  ggplot(aes(model_2b_preds, model_2b_resids)) + 
  geom_point() + 
  stat_smooth()
```
###iii.No Perfect Collinearity
First, check if any variables were dropped by R. No coefficient is missing, which mean there is no perfect collinearity.

```{r}
model_2b$coefficients
```
In addition, all the variance inflation factors are less than 4, which doesn't indicate the exisitence of collinearity. 
```{r}
vif(model_2b)
```
###iv.Homoskedastic Errors
32 is New Mexico, 33 is New York, 35 is North Dakota

```{r, echo=FALSE}
plot(model_2b, which=3)
```
###v.Normally Distributed Errors
Check the normality of error distribution based on the histogram of residuals and the qqplot.

```{r, echo=FALSE}
plot_one <- df %>% 
  ggplot(aes(x = model_2b_resids)) + 
  stat_bin(bins = 25)
  geom_histogram()

plot_two <- df %>% 
  ggplot(aes(sample = model_2b_resids)) + 
  stat_qq() + stat_qq_line()

plot_one / plot_two
```

```{r, echo=FALSE}
df %>% 
  select(case_rate_100k, senior_per_100k, log_population_density, noneld_pre_existing_per_100k, poverty_per_100k, white_per_100k) %>% 
  GGally::ggpairs()
```
### Cook Distance

```{r, echo =FALSE}
plot(model_2b, which = 5)
```